```{r echo = F}
# Strings to match samples and create labels
interaction <- "{{interaction}}"
sect_title <- "{{name}}"
if(interaction){
  de_test <- contrasts[[sect_title]][[1]]
  comparison <- contrasts[[sect_title]][[2]]

  comparison <- unlist(strsplit(comparison, split = ":"))
  control <- comparison[2]
  treatment <- comparison[1]
} else {
  control <- contrasts[[sect_title]][2]
  treatment <- contrasts[[sect_title]][1]

}
```

## **`r sect_title`**

```{r run-de-{{name}}}
# Extract DE genes
if(interaction){
  sig_genes <- get_de(dds, de_test = de_test, interaction = interaction,
                      var1 = treatment, var2 = control,
                      write_csv = T, p_value = params$DE_alpha,
                      lfc = 0, output_dir = output_directory,
                      lfc_shrink = T, save_name = sect_title)

  sig_genes_high_lfc <- get_de(dds, de_test = de_test,
                               interaction = interaction, write_csv = F,
                               p_value = params$DE_alpha,
                               lfc = params$DE_lfc,
                               lfc_shrink = T)
  
  sig_genes_all <- get_de(dds, de_test = de_test, interaction = interaction,
                          var1 = treatment, var2 = control,
                          write_csv = T, p_value = 1,
                          lfc = 0, output_dir = output_directory,
                          lfc_shrink = T, save_name = paste0(sect_title, "_all"))
} else {
  sig_genes <- get_de(dds, column = params$sample_column, var1 = treatment,
                      var2 = control, interaction = interaction,
                      write_csv = T, p_value = params$DE_alpha,
                      lfc = 0, output_dir = output_directory,
                      lfc_shrink = T, save_name = sect_title)

  sig_genes_high_lfc <- get_de(dds, column = params$sample_column,
                               var1 = control, var2 = treatment,
                               interaction = interaction, write_csv = F,
                               p_value = params$DE_alpha,
                               lfc = params$DE_lfc,
                               lfc_shrink = T)
  all_genes_all <- get_de(dds, column = params$sample_column, var1 = treatment,
                          var2 = control, interaction = interaction,
                          write_csv = T, p_value = 1,
                          lfc = 0, output_dir = output_directory,
                          lfc_shrink = T, save_name = paste0(sect_title, "_all"))
}
DE_genes_n <- nrow(sig_genes$DE_genes)
```

### DEseq2 analysis
**`r format(DE_genes_n, big.mark = ",")`** differentially expressed genes were identified using the `DESeq2` package.

Differentially expressed genes are labeled on the MA plot in blue.
```{r, fig.width = 6, fig.height = 4}
plotMA(sig_genes$DE_df, main = sect_title)
```

### Heatmaps
Heatmap of DE genes across all samples
```{r heatmaps-{{name}}, fig.width = 6, fig.height = 4}
heatmap_dir <- file.path(output_directory, "images", "heatmaps")
heatmap <- make_heatmap(dds = dds,
                        vsd = vsd,
                        de_genes = sig_genes$DE_genes,
                        treatment = treatment,
                        control = control,
                        group = params$sample_column,
                        print_genenames = FALSE,
                        cluster_cols = FALSE,
                        save_heatmap = TRUE,
                        output_dir = heatmap_dir,
                        color_test = genotype_colors,
                        save_name = sect_title)
print(heatmap)

```

Heatmap of DE genes across only the samples used in the comparison

```{r heatmaps2-{{name}}, fig.width = 6, fig.height = 4}
heatmap_dir <- file.path(output_directory, "images", "heatmaps")
heatmap <- make_heatmap(dds = dds,
                        vsd = vsd,
                        de_genes = sig_genes$DE_genes,
                        treatment = treatment,
                        control = control,
                        group = params$sample_column,
                        print_genenames = FALSE,
                        cluster_cols = FALSE,
                        save_heatmap = FALSE,
                        plot_groups = c(treatment, control),
                        color_test = genotype_colors)
print(heatmap)

```

### Gene Set Enrichment

For Gene Set Enrichment, I used `gprofiler2` It uses a hypergeometric test to determine overrepresentation of genes from different categories.

```{r gse-{{name}}, fig.height = 30, fig.width = 12, eval = FALSE}
all_gene_ens <- sub(".*_ENSMUSG", "ENSMUSG", rownames(dds))

evcode <- TRUE

result <- run_gprofiler_all(gene_table = sig_genes$DE_genes,
                            pos_name = treatment,
                            neg_name = control,
                            custom_bg = all_gene_ens,
                            save_dir = file.path(output_directory, "GSE_files"),
                            plot_dir = file.path(output_directory, "images",
                                                 "GSE"),
                            save_name = sect_title,
                            evcodes = evcode)

plots <- result$plots
gost_table <- result[[paste0(treatment, "_vs_", control)]]$result

save_gost <- c("GO:BP", "GO:CC", "GO:MF", "KEGG")

gost_file <- openxlsx::createWorkbook()

invisible(lapply(save_gost, function(x){
  save_table <- gost_table %>%
    dplyr::filter(source == x)
    
  if(evcode){
    save_table <- save_table %>%
      dplyr::select(!c(intersection, evidence_codes))

  }
  sheetname <- sub(":", "_", x)
  openxlsx::addWorksheet(wb = gost_file, sheetName = sheetname)
  openxlsx::writeData(wb = gost_file, sheet = sheetname, x = save_table)
}))

openxlsx::saveWorkbook(wb = gost_file,
                       file = file.path(output_directory, "GSE_files", 
                                     paste0(sect_title, "_GSE.xlsx")),
                       overwrite = TRUE)

plot_grid(plots$GOBP, plots$GOMF, plots$GOCC, plots$KEGG,
          nrow = 4,
          ncol = 1,
          align = "v",
          axis = "tb")

```


### Gene Set Enrichment Separating Up and Down Regulation

For Gene Set Enrichment, I used `gprofiler2` It uses a hypergeometric test to determine overrepresentation of genes from different categories.

```{r gse-split-{{name}}, fig.height = 30, fig.width = 12, eval = FALSE}
all_gene_ens <- sub(".*_ENSMUSG", "ENSMUSG", rownames(dds))

ifelse(!dir.exists(file.path(params$output_dir, "GSE_files_separated")),
       dir.create(file.path(params$output_dir, "GSE_files_separated")),
       FALSE)

result <- run_gprofiler(gene_table = sig_genes$DE_genes,
                        pos_name = treatment,
                        neg_name = control,
                        custom_bg = all_gene_ens,
                        save_dir = file.path(params$output_dir,
                                             "GSE_files_separated"),
                        plot_dir = file.path(params$output_dir, "images",
                                             "GSE"),
                        evcodes = FALSE)

plots <- result$plots
gost_table_trt <- result[[treatment]]$result
gost_table_ctl <- result[[control]]$result


save_gost <- c("GO:BP", "GO:CC", "GO:MF", "KEGG", "WP")

gost_file_trt <- openxlsx::createWorkbook()

gost_file_ctl <- openxlsx::createWorkbook()

add_gost_data <- function(gost_table, gost_file, source_test,
                          evcode = FALSE){
  save_table <- gost_table %>%
    dplyr::filter(source == source_test)
  
  if(evcode){
    save_table <- save_table %>%
          dplyr::select(!c(intersection, evidence_codes))
  }

  sheetname <- sub(":", "_", source_test)
  openxlsx::addWorksheet(wb = gost_file, sheetName = sheetname)
  openxlsx::writeData(wb = gost_file, sheet = sheetname, x = save_table)
}

invisible(lapply(save_gost, function(x){
  add_gost_data(gost_table_trt, gost_file_trt, x, evcode = FALSE)
  add_gost_data(gost_table_ctl, gost_file_ctl, x, evcode = FALSE)
}))

openxlsx::saveWorkbook(wb = gost_file_trt,
                       file = file.path(params$output_dir, "GSE_files_separated", 
                                     paste0(sect_title, "_", treatment, 
                                     "_upregulated_genes_GSE.xlsx")),
                       overwrite = TRUE)

openxlsx::saveWorkbook(wb = gost_file_ctl,
                       file = file.path(params$output_dir, "GSE_files_separated", 
                                     paste0(sect_title, "_", control, 
                                     "_upregulated_genes_GSE.xlsx")),
                       overwrite = TRUE)

plot_grid(plots$C_GOBP, plots$T_GOBP,
          plots$C_GOMF, plots$T_GOMF,
          plots$C_GOCC, plots$T_GOCC,
          plots$C_KEGG, plots$T_KEGG,
          nrow = 8,
          ncol = 1,
          align = "v",
          axis = "tb")

```