```{r echo = F}

# Strings to match samples and create labels
#gene <- "{{.x}}"
sample <- "{{sample}}"
sect_title <- sample

if(!exists("extra_pound_quality")){
  extra_pound_quality <- ""
}

sample_info <- sample_list[[sample]]

```

###`r extra_pound_quality` **`r sect_title`** {.tabset}

####`r extra_pound_quality` RNA

```{r {{sample}}-quality, fig.height=4, fig.width=6}
if(sample_info$type == "single"){
  unprocessed_object <- readRDS(sample_info$seurat_unprocessed_path)
} else {
  unprocessed_object <- sample_info$seurat_object
}

Idents(unprocessed_object) <- "orig.ident"
rna_qual <- VlnPlot(unprocessed_object,
                    features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),
                    ncol = 3)
rna_qual
```

```{r}
median_counts <- median(unprocessed_object$nCount_RNA)
median_features <- median(unprocessed_object$nFeature_RNA)
cell_number <- nrow(unprocessed_object[[]])
```

#####`r extra_pound_quality` Metrics

* The median counts per cell is `r median_counts`
* The median genes per cell is `r median_features`
* The total number of cells before filtering is `r cell_number`

####`r extra_pound_quality` ADT

```{r {{sample}}-adt-qual, fig.height=4, fig.width=4}
adt_qual <-   adt_qual <- VlnPlot(unprocessed_object,
                                  features = c("nCount_ADT", "nFeature_ADT"))

adt_qual
```

```{r}
median_counts_adt <- median(unprocessed_object$nCount_ADT)
median_features_adt <- median(unprocessed_object$nFeature_ADT)
```

#####`r extra_pound_quality` Metrics

* The median ADT counts per cell is `r median_counts_adt`
* The median ADTs per cell is `r median_features_adt`
